Backward Population Synthesis: Mapping the Evolutionary History of Gravitational-Wave Progenitors

### 2022-11-29 src/tex/ms.tex
Add symbols to table one

\begin{table*}[hbt!]
    \begin{center}
    \begin{tabular}{ l l l }
    \hline
    \hline
    Parameters &  Description & Optimization range\\
    \hline
    \hline
    Observables &\ &\  \\
    \hline
    \hline
    $m_{\rm 1, GW}$ & Primary mass of the GW event & NA \\
    $m_{\rm 2, GW}$ & Secondary mass of the GW event  & NA\\
    \hline
    \hline
    Progenitor parameters &\ &\  \\
    \hline
    \hline
    $m_{\rm 1, ZAMS}$ & Primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
    $m_{\rm 2, ZAMS}$ & Secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
    $t_{\rm orb}$ & Orbital period at ZAMS & $[5,5000]$ days\\
    $e$ & Eccentricity at ZAMS & $[0,1]$\\
    $Z$ & Metallicity at ZAMS & $[10^{-4},2\times10^{-2}]$\\
    $z$ & Redshift at ZAMS & NA\\
    \hline
    \hline
    hyperparameters &\ &\ \\
    \hline
    \hline

@@

We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
We use the \texttt{Overall\_posterior} posterior samples for GW150914,
publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
The spin of BHs formed in isolated binaries depends strongly on the angular
momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
therefore we only consider the two component masses in the observable space in both the
root-finding stage and the MCMC stage. For progenitor parameters, we characterize each guess with
five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
obtained through the root-finding process because the formation redshift does not intrinsically affect
the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
time to merger to be less than $13.7\,\rm{Gyr}$.
Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the
Planck 2018 cosmological model \citep{Planck2018}.
The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
by reweighting the posterior distribution according to a particular SFR model,
and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
For hyperparameters, we choose parameters that strongly affect
binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
the critical mass ratio which determines whether Roche-overflow mass transfer
proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
employed for CO natal kicks, $\sigma$. For systems which begin a common envelope 
with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
stars merge since the donor has likely not formed a strong core-envelope boundary 
\citep{Ivanova2004, Belczynski2008}.

@@

\begin{figure*}[h]
    \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
    \caption{The posterior for GW150914 in both the progenitor parameter space
    and hyperparameter space. $M_1$ and $M_2$ are the progenitors' masses.
    $\log{Z}$ is log metallicity at ZAMS. $z$ is the redshift at ZAMS. $\alpha$
    is the common envelope efficiency. $f_{\rm acc}$ is the fraction of mass
    accreted during stable mass transfer. $q_{\rm crit, 3}$ is the critical mass
    ratio on the Hertzsprung Gap. The contours correspond to the $68\%$ and
    $95\%$ confidence intervals. Note that the redshift is not fitted during the
    root finding process or the MCMC process. Once we find the evolutionary
    parameters, we add the delay time to the lookback time of the observed
    posterior sample, then from the total lookback time we can compute the
    redshift at ZAMS. We highlight three panels in the corner plots to show the
    fine structure of the set of posterior samples in the evolutionary parameter
    space. We also color the posterior in a particular panel according to the
    type of parameters involved in the corner plot. Blue denotes panels that
    include only progenitor parameters, green denotes panels that include a mix
    of progenitor parameters and hyperparameters, and orange denotes panels that
    include only hyperparameters. }
    \label{fig:GW150914_posterior}
    \script{figure1_cornerPlot.py}
\end{figure*}

@@

\begin{figure*}
    \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like
     mergers for three variations of $\alpha$ with a fixed set of parameter
     assumptions matching those of A21 (first three columns) and for binaries
     which produce GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and
     $q_{crit,3}$ are allowed to vary, (fourth column).}
    \label{fig:compare_fixed_variable}
    \script{figure2_varyAlphaFacc.py}
\end{figure*}

@@

\begin{figure}
\includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
\caption{Reprojecting the posterior in evolutionary parameter space of GW150914
to observable space. The contours correspond to the $68\%$ and $95\%$ confidence
intervals. The blue contour is the reprojected posterior after the root-finding
procedure. The orange contour is the reprojected posterior after the MCMC
procedure. The green contour is the posterior plotted using the original LVK
posterior samples. }
\label{fig:GW150914_reprojection}
\script{figure3_sanityCheck.py}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
\caption{ The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
    most of the events in GWTC-3. Note that $m_{1,{\rm GW}}$ here refers to the
    reprojected posterior instead of the posterior released by LIGO. Each
    contour is the $68\%$ credible interval of the posterior density for a
    particular event. At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair
    instability supernova mechanism prevents \texttt{COSMIC} from producing
    events that are more massive than this cutoff. Therefore, events with the
    majority of posterior support above this cutoff are not compatible with
    \texttt{COSMIC}, hence have a large KL divergence and are excluded from this
    figure. On the low mass end, neutron star binaries or neutron star-black
    hole binaries are subjected to randomness induced by the natal kick, also
    resulting in a larger and fluctuating KL divergences, and therefore are also
    excluded from the analysis. }
\label{fig:GWTC-3_f_acc_mass}
\script{figure4_population.py}
\end{figure}

>>>>

\begin{table*}[hbt!]
    \begin{center}
    \begin{tabular}{ l l l }
    \hline
    \hline
    Parameters &  Description & Optimization range\\
    \hline
    \hline
    Observables $\bm{\theta}$ &\ &\  \\
    \hline
    \hline
    $m_{\rm 1, GW}$ & Primary mass of the GW event & NA \\
    $m_{\rm 2, GW}$ & Secondary mass of the GW event  & NA\\
    \hline
    \hline
    Progenitor parameters $\bm{\theta'}$&\ &\  \\
    \hline
    \hline
    $m_{\rm 1, ZAMS}$ & Primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
    $m_{\rm 2, ZAMS}$ & Secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
    $t_{\rm orb}$ & Orbital period at ZAMS & $[5,5000]$ days\\
    $e$ & Eccentricity at ZAMS & $[0,1]$\\
    $Z$ & Metallicity at ZAMS & $[10^{-4},2\times10^{-2}]$\\
    $z$ & Redshift at ZAMS & NA\\
    \hline
    \hline
    Hyperparameters $\bm{\lambda    }$ &\ &\ \\
    \hline
    \hline

@@

We apply the method described in section \ref{sec:method} to the first GW event GW150914 \citep{GW150914}.
We use the \texttt{Overall\_posterior} posterior samples for GW150914,
publicly available from the Gravitational wave Open Science Center \citep{LIGOScientific:2019lzm}. 
The parameters involved in the analysis and the range we allow for the inference are tabulated in table \ref{tab:parameters}.
The spin of BHs formed in isolated binaries depends strongly on the angular
momentum transport in massive stars which is still highly uncertain (see however \citet{Fuller2019, Bavera2020}),
therefore we only consider the two component masses in the observable space $\bm{\theta}$ in both the
root-finding stage and the MCMC stage. For progenitor parameters $\bm{\theta'}$, we characterize each guess with
five parameters: the two component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period $t_{\rm orb}$,
the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that the progenitor formation redshift is not
obtained through the root-finding process because the formation redshift does not intrinsically affect
the evolution of the binary. Instead, we compute the time it takes for the binary to evolve from
formation to the merger and add this time to the lookback time when a GW event is observed and limit the total
time to merger to be less than $13.7\,\rm{Gyr}$.
Using the total lookback time, we can compute the redshift of the binary at ZAMS formation following the
Planck 2018 cosmological model \citep{Planck2018}.
The main benefit of post-processing the redshift in this way is that we do not rely on a particular assumption
of star formation rate (SFR) distribution when we solve for the posterior distribution in redshift.
This means we can put the prior on formation redshift (or in general, metallicity-redshfit distribution) in
by reweighting the posterior distribution according to a particular SFR model,
and therefore do not have to rerun the inference every time we change the SFR model. This is especially helpful
since SFR models are still highly uncertain but strongly impact the local merger rates of BBHs \citep[e.g. ][]{Broekgaarden2021}.
For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
the critical mass ratio which determines whether Roche-overflow mass transfer
proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
employed for CO natal kicks, $\sigma$. For systems which begin a common envelope 
with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
stars merge since the donor has likely not formed a strong core-envelope boundary 
\citep{Ivanova2004, Belczynski2008}.

@@

% \begin{figure*}[h]
%     \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
%     \caption{The posterior for GW150914 in both the progenitor parameter space
%     and hyperparameter space. $M_1$ and $M_2$ are the progenitors' masses.
%     $\log{Z}$ is log metallicity at ZAMS. $z$ is the redshift at ZAMS. $\alpha$
%     is the common envelope efficiency. $f_{\rm acc}$ is the fraction of mass
%     accreted during stable mass transfer. $q_{\rm crit, 3}$ is the critical mass
%     ratio on the Hertzsprung Gap. The contours correspond to the $68\%$ and
%     $95\%$ confidence intervals. Note that the redshift is not fitted during the
%     root finding process or the MCMC process. Once we find the evolutionary
%     parameters, we add the delay time to the lookback time of the observed
%     posterior sample, then from the total lookback time we can compute the
%     redshift at ZAMS. We highlight three panels in the corner plots to show the
%     fine structure of the set of posterior samples in the evolutionary parameter
%     space. We also color the posterior in a particular panel according to the
%     type of parameters involved in the corner plot. Blue denotes panels that
%     include only progenitor parameters, green denotes panels that include a mix
%     of progenitor parameters and hyperparameters, and orange denotes panels that
%     include only hyperparameters. }
%     \label{fig:GW150914_posterior}
%     \script{figure1_cornerPlot.py}
% \end{figure*}

@@

% \begin{figure*}
%     \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
%     \caption{Comparison of ZAMS masses for binaries which produce GW150914-like
%      mergers for three variations of $\alpha$ with a fixed set of parameter
%      assumptions matching those of A21 (first three columns) and for binaries
%      which produce GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and
%      $q_{crit,3}$ are allowed to vary, (fourth column).}
%     \label{fig:compare_fixed_variable}
%     \script{figure2_varyAlphaFacc.py}
% \end{figure*}

@@

% \begin{figure}
% \includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
% \caption{Reprojecting the posterior in evolutionary parameter space of GW150914
% to observable space. The contours correspond to the $68\%$ and $95\%$ confidence
% intervals. The blue contour is the reprojected posterior after the root-finding
% procedure. The orange contour is the reprojected posterior after the MCMC
% procedure. The green contour is the posterior plotted using the original LVK
% posterior samples. }
% \label{fig:GW150914_reprojection}
% \script{figure3_sanityCheck.py}
% \end{figure}

@@

% \begin{figure}
% \includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
% \caption{ The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
%     most of the events in GWTC-3. Note that $m_{1,{\rm GW}}$ here refers to the
%     reprojected posterior instead of the posterior released by LIGO. Each
%     contour is the $68\%$ credible interval of the posterior density for a
%     particular event. At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair
%     instability supernova mechanism prevents \texttt{COSMIC} from producing
%     events that are more massive than this cutoff. Therefore, events with the
%     majority of posterior support above this cutoff are not compatible with
%     \texttt{COSMIC}, hence have a large KL divergence and are excluded from this
%     figure. On the low mass end, neutron star binaries or neutron star-black
%     hole binaries are subjected to randomness induced by the natal kick, also
%     resulting in a larger and fluctuating KL divergences, and therefore are also
%     excluded from the analysis. }
% \label{fig:GWTC-3_f_acc_mass}
% \script{figure4_population.py}
% \end{figure}
